---
title: "Fig. 2 - FMT vs. Memory"
author: "Tom Morin"
date: "2025-04-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = FALSE}
# Load Libraries
library(tidyverse) # v2.0.0
library(ggpubr) # v0.6.0
library(lme4) # v1.1.32
library(lmerTest) # v3.1.3
library(rstatix) # v0.7.2
library(emmeans) # v1.10.0
library(lsmeans) #2.30.2

# Set colorscales
biomarker_group_cscale <- scale_color_manual(values = c("#549e3e", "#323690", "#B6357B"))
biomarker_group_fscale <- scale_fill_manual(values = c("#549e3e", "#323690", "#B6357B"))

# Read in data
data_BCD <- read_csv('data_fig2bcd.csv')
data_E <- read_csv('data_fig2e.csv')

data_E$biomarker_group <- factor(data_E$biomarker_group,
                                 levels = c('Younger\nAdults',
                                            'Normal\nAging',
                                            'Preclinical\nAD'))
```

## Plot Panel B: FMT vs. Age Group
```{r, echo = FALSE}
# Panel B
figB <- ggbarplot(data_BCD, 
          x = 'age_group', 
          y = 'fmt_ki_vst', 
          fill = 'age_group', 
          color = 'black', 
          alpha = 0.5,
          palette = c("#549e3e", "#643f95"), 
          add = c('mean_se'), 
          legend = 'none',
          order = c('Younger\nAdults', 'Older\nAdults')) +
          geom_jitter(aes(color = age_group, 
                          fill = age_group, 
                          alpha = 0.5),
                          position = position_jitterdodge(0.7))

age.stat.test <- data_BCD %>% t_test(fmt_ki_vst ~ age_group, detailed = TRUE)
age.stat.test.p <- age.stat.test %>%
  add_xy_position(x = 'age_group', fun = 'mean_se', dodge = 0.8)
age.stat.test.p <- age.stat.test.p %>% mutate(sig = case_when(p < 0.05 ~ '*',
                                                              p >= 0.05 ~ ''))

figB <- figB + stat_pvalue_manual(age.stat.test.p, label = 'sig', tip.length = 0.03, 
                       step.increase = 0.08, y.position = 0.03,
                       size = 8, bracket_size = 1, hide.ns = T) + 
    xlab('') + 
    ylab('[18F]FMT Ki\n Ventral Striatum)') +
    scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
    theme_classic() +
    theme(panel.background = element_blank(),
        axis.line = element_line(color = 'black'),
        legend.position = 'none', 
        axis.text.x = element_text(color = 'black', size = 10),
        axis.text.y = element_text(color = 'black', size = 10),
        axis.title.x = element_text(color = 'black', size = 12),
        axis.title.y = element_text(color = 'black', size = 12),
        legend.title = element_blank())
figB
```

## Does FMT signal differ across age groups? Across biomarker groups?
```{r}
# Test for difference in FMT signal based on age group
age.stat.test <- data_BCD %>% 
    t_test(fmt_ki_vst ~ age_group, detailed = TRUE)
age.stat.test.p <- age.stat.test %>% mutate(sig = case_when(p < 0.05 ~ '*',
                                                        p >= 0.05 ~ ''))
age.stat.test

# Test for difference in FMT signal based on preclinical AD status
biomarker.stat.test <- data_BCD %>% 
    t_test(fmt_ki_vst ~ biomarker_group, detailed = TRUE)
biomarker.stat.test.p <- biomarker.stat.test %>% mutate(sig = case_when(p < 0.05 ~ '*',
                                                        p >= 0.05 ~ ''))
biomarker.stat.test
```

## Plot Panel C: FMT vs. PiB
```{r, echo = FALSE}
# Panel C: Plot FMT vs. FTP
figC <- data_BCD %>% filter(age_group == 'Older\nAdults') %>%
  ggplot(aes(y = pib_dvr, x = fmt_ki_vst)) +
  geom_point(color = "#643f95", fill = "#643f95", alpha = 0.5) +
  geom_smooth(method = 'lm', color = "#643f95", fill = "#643f95",
              linewidth = 2, level = 0.9) +
  theme_classic() +
  ylab('[18F]PiB DVR\n(Cortex)') +
  xlab('[18F]FMT Ki\n(Ventral Striatum)') +
  theme(panel.background = element_blank(),
      axis.line = element_line(color = 'black'),
      legend.position = 'bottom', 
      axis.text.x = element_text(color = 'black', size = 10),
      axis.text.y = element_text(color = 'black', size = 10),
      axis.title.x = element_text(color = 'black', size = 12),
      axis.title.y = element_text(color = 'black', size = 12),
      legend.title = element_blank())

figC
```

## Is FMT signal related to PiB signal?
```{r, echo = FALSE}
mod <- lm(fmt_ki_vst ~ pib_dvr + age + sex + yrs_edu,
          data = data_BCD)
summary(mod)
```

## Plot Panel D: FMT vs. Tau
```{r, echo = FALSE}
# Panel D: Plot FMT vs. FTP
figD <- data_BCD %>% filter(age_group == 'Older\nAdults') %>%
  ggplot(aes(y = ftp_braak34_suvr, x = fmt_ki_vst)) +
  geom_point(color = "#643f95", fill = "#643f95", alpha = 0.5) +
  geom_smooth(method = 'lm', color = "#643f95", fill = "#643f95",
              linewidth = 2, level = 0.9) +
  theme_classic() +
  ylab('[18F]FTP SUVR\n(Braak 3/4 Regions)') +
  xlab('[18F]FMT Ki\n(Ventral Striatum)') +
  theme(panel.background = element_blank(),
      axis.line = element_line(color = 'black'),
      legend.position = 'bottom', 
      axis.text.x = element_text(color = 'black', size = 10),
      axis.text.y = element_text(color = 'black', size = 10),
      axis.title.x = element_text(color = 'black', size = 12),
      axis.title.y = element_text(color = 'black', size = 12),
      legend.title = element_blank())

figD
```

## Is FMT signal related to FTP signal?
```{r, echo = FALSE}
mod <- lm(fmt_ki_vst ~ ftp_braak34_suvr + age + sex + yrs_edu,
          data = data_BCD)
summary(mod)
```

## Plot Panel E: FMT signal vs. Memory, Faceted by Valence and Biomarker Group
```{r, echo = FALSE}
# Panel E: Plot FMT by d' across each valence
figE <- data_E %>% filter(!is.na(valence)) %>%
  ggplot(aes(x = fmt_ki_vst, y = dprime, group = biomarker_group,
             color = biomarker_group)) +
  geom_point(alpha = 0.5) +
  geom_smooth(aes(color = biomarker_group, fill = biomarker_group),
                  method = 'lm', linewidth = 2, level = 0.9) +
  biomarker_group_cscale +
  biomarker_group_fscale +
  theme_classic() +
  xlab(bquote('[18F]FMT Ki\n(Ventral Striatum)')) +
  ylab('d\'') +
  facet_wrap(.~valence, nrow = 1) +
  theme(panel.background = element_blank(),
      axis.line = element_line(color = 'black'),
      legend.position = 'bottom', 
      axis.text.x = element_text(color = 'black', size = 8),
      axis.text.y = element_text(color = 'black', size = 8),
      axis.title.x = element_text(color = 'black', size = 12),
      axis.title.y = element_text(color = 'black', size = 12),
      legend.title = element_blank())

figE
```

## Is FMT signal related to memory performance? Does this differ by valence or biomarker group?

### Initial analysis in Young Adults only
```{r, echo = FALSE}
lme0 <- lmer(dprime ~ fmt_ki_vst*valence + (1 | subject),
             data_E %>% filter(age_group == 'Younger\nAdults'))
summary(lme0)
```

### Initial analysis in Older Adults only
```{r, echo = FALSE}
lme0 <- lmer(dprime ~ fmt_ki_vst*valence + (1 | subject),
             data_E %>% filter(age_group == 'Older\nAdults'))
summary(lme0)
```

### Test for effects of FMT, biomarker group, valence, and interactions on memory
```{r, echo = FALSE}
# Stats for Panel E:
lme1 <- lmer(dprime ~ fmt_ki_vst*biomarker_group + valence*biomarker_group + (1 | subject),
             data_E)
summary(lme1)

print('============================== LME 1 POSTHOC TESTS ==============================')
slopes <- lstrends(lme1, ~ biomarker_group, var = 'fmt_ki_vst')
summary(slopes, infer=TRUE)
```

### Repeat model, accounting for learning performance
```{r, echo = FALSE}
lme2 <- lmer(dprime ~ fmt_ki_vst*biomarker_group + valence*biomarker_group + learning + (1 | subject),
             data_E)
summary(lme2)

print('============================== LME 2 POSTHOC TESTS ==============================')
slopes <- lstrends(lme2, ~ biomarker_group, var = 'fmt_ki_vst')
summary(slopes, infer=TRUE)
```

## Plot complete Figure 2
```{r, echo = FALSE}
# Combined Panels
fmt_fig <- ggarrange(ggarrange(figB, figB, labels = c('A', 'B'), ncol = 2), 
                     ggarrange(figC, figD, labels = c('C', 'D'), ncol = 2),
          figE, nrow = 3, labels = c('', '', 'E'), heights = c(1, 1, 1.3))

ggsave(fmt_fig, filename = 'fmt_fig.pdf', width = 5, height = 8, units = 'in')
fmt_fig
```