---
title: "Positivity Effect"
author: "Tom Morin"
date: "2024-07-31"
output: html_document
---

```{r, echo = FALSE}
# Load Libraries
library(tidyverse)
library(ggpubr)
library(rstatix)
library(lme4)
library(lmerTest)
library(lsmeans)

# Set colorscales
age_group_cscale <- scale_color_manual(values = c("#549e3e", "#643f95"))
age_group_fscale <- scale_fill_manual(values = c("#549e3e", "#643f95"))

jack_status_cscale <- scale_color_manual(values = c("#549e3e", "#323690", "#B6357B"))
jack_status_fscale <- scale_fill_manual(values = c("#549e3e", "#323690", "#B6357B"))

healthy_cscale <- scale_color_manual(values = c("#549e3e", "#323690"))
healthy_fscale <- scale_fill_manual(values = c("#549e3e", "#323690"))

npg_pal <- get_palette(palette = 'npg', 3)
npg_RL <- c(npg_pal[1], npg_pal[3])

# Read in data
data_A <- read_csv('data_fig4a.csv')
data_A$age_group <- factor(data_A$age_group, levels = c('Younger\nAdults',
                                                        'Older\nAdults'))
data_A$biomarker_group <- factor(data_A$biomarker_group, levels = c('Younger\nAdults',
                                                                    'Normal\nAging',
                                                                    'Preclinical\nAD'))
data_BC <- read_csv('data_fig4bc.csv')
data_BC$age_group <- factor(data_BC$age_group, levels = c('Younger\nAdults',
                                                        'Older\nAdults'))
data_BC$biomarker_group <- factor(data_BC$biomarker_group, levels = c('Younger\nAdults',
                                                                    'Normal\nAging',
                                                                    'Preclinical\nAD'))
```

## Plot Panel A: Age-related positivity effect
```{r, echo = FALSE}
# PANEL A: Positivity Effect
age_comparisons = list(c('Older\nAdults', 'Younger\nAdults'))
biomarker_group_comparisons <- list(c('Younger\nAdults', 'Normal\nAging'),
                                c('Younger\nAdults', 'Preclinical\nAD'))

figA <- data_A %>%
  ggplot(aes(x = age_group, y = pos_index, color=age_group, fill=age_group)) +
  geom_hline(yintercept = 0) +
  geom_violin(alpha = 0.5, linewidth = 1) +
  geom_jitter(alpha = 0.5, size = 2, position = position_jitterdodge(0.2)) +
  stat_summary(geom = 'pointrange', fun.data = 'mean_cl_boot', size = 1, linewidth = 2) +
  age_group_cscale +
  age_group_fscale +
  xlab('') +
  ylab(bquote(atop('Positivity Index', (H[Rew] - H[Loss])/(H[Rew] + H[Loss])))) +
  theme_classic() +
  stat_compare_means(comparisons = age_comparisons, 
                     label = 'p.signif', 
                     size = 5,
                     hide.ns = T) +
  theme(panel.background = element_blank(),
        axis.line = element_line(color = 'black'),
        legend.position = 'none', 
        axis.text.x = element_text(color = 'black', size = 8),
        axis.text.y = element_text(color = 'black', size = 8),
        axis.title.x = element_text(color = 'black', size = 10),
        axis.title.y = element_text(color = 'black', size = 10),
        legend.title = element_blank()
    
  )

figA
```

# T-test and linear models: Age group difference in positivity index
```{r, echo = FALSE}
t_test(data_A, pos_index ~ age_group)

# PANEL C STATS MODELS:
mod <- lm(pos_index ~ biomarker_group, data = data_A)
summary(mod)

mod_learning <- lm(pos_index ~ biomarker_group + learning_RewVsLoss, 
                   data = data_A)
summary(mod_learning)  
```

## Plot Panel B: Interaction between biomarker group and FMT on Positivity Index
```{r, echo = FALSE}
figB <- data_BC %>%
  ggplot(aes(x = fmt_ki_vst, y = pos_index, group = biomarker_group,
             color = biomarker_group, fill = biomarker_group)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = 'lm', level = 0.9, linewidth = 2) +
  jack_status_cscale +
  jack_status_fscale +
  theme_classic() +
  xlab(bquote('[18F]FMT Ki\n(Ventral Striatum)')) +
  ylab(bquote(atop('Positivity Index', (H[Rew] - H[Loss])/(H[Rew] + H[Loss])))) +
  theme(panel.background = element_blank(),
      axis.line = element_line(color = 'black'),
      legend.position = 'none', 
      axis.text.x = element_text(color = 'black', size = 8),
      axis.text.y = element_text(color = 'black', size = 8),
      axis.title.x = element_text(color = 'black', size = 10),
      axis.title.y = element_text(color = 'black', size = 10),
      legend.title = element_blank())

figB
```

## Linear Model: Interaction between Biomarker Group and FMT on Positivity Index
```{r, echo = FALSE}
# Stats for Panel B

# Simplest model
lm1 <- lm(pos_index ~ fmt_ki_vst*biomarker_group, data = data_BC)
summary(lm1)

lm1 <- lm(pos_index ~ fmt_ki_vst*age_group, data = data_BC)
summary(lm1)
slopes <- lstrends(lm1, ~age_group, var = 'fmt_ki_vst')
summary(slopes)
pairs(slopes)
```

## Repeate Analysis, accounting for learning performance
```{r, echo = FALSE}
# Model accounting for learning accuracy
lm2 <- lm(pos_index ~ fmt_ki_vst*biomarker_group + learning_RewVsLoss,
          data = data_BC)
summary(lm2)

```

## Plot Panel C: Quadratic relationship?
```{r, echo = FALSE}
figC <- data_BC %>%
  ggplot(aes(x = fmt_ki_vst, y = pos_index)) +
  geom_smooth(method = 'lm', formula = y ~ x + I(x^2), linewidth = 2, color = 'gray') +
  geom_point(aes(color = biomarker_group, fill = biomarker_group), alpha = 0.5) +
  geom_smooth(method = 'lm', aes(color = biomarker_group), linewidth = 1.5, 
              linetype = 'dashed', se = F) +
  jack_status_cscale +
  jack_status_fscale +
  theme_classic() +
  xlab(bquote('[18F]FMT Ki\n(Ventral Striatum)')) +
  ylab(bquote(atop('Positivity Index', (H[Rew] - H[Loss])/(H[Rew] + H[Loss])))) +
  theme(panel.background = element_blank(),
      axis.line = element_line(color = 'black'),
      legend.position = 'none', 
      axis.text.x = element_text(color = 'black', size = 8),
      axis.text.y = element_text(color = 'black', size = 8),
      axis.title.x = element_text(color = 'black', size = 10),
      axis.title.y = element_text(color = 'black', size = 10),
      legend.title = element_blank())

figC
```

## Quadratic Regression: FMT signal vs. Positivity Index
```{r, echo = FALSE}
quad_all <- lm(pos_index ~ fmt_ki_vst + I(fmt_ki_vst^2), 
                   data = data_BC %>% filter(!is.na(fmt_ki_vst)))
summary(quad_all)
```

## Plot entire Figure 4
```{r, echo = FALSE}
blank <- ggplot() + theme_void()

pos_fig <- ggarrange(
             ggarrange(figA, figB, blank, nrow = 1,
                       widths = c(1, 1, 0), labels = c('A', 'B', '')),
             ggarrange(blank, figC, blank, nrow = 1,
                       widths = c(0.5, 1, 0.5), labels = c('', 'C', '')),
            nrow = 2
)

ggsave(pos_fig, filename = 'fig4.pdf', width = 6, height = 6, units = 'in')
pos_fig
```