---
title: "Fig s1-2cdef - Tau Burden vs. Memory/Learning"
author: "Tom Morin"
date: "2025-12-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load Libraries
library(tidyverse) # v2.0.0
library(ggpubr) # v0.6.0
library(rstatix) # v0.7.2
library(lme4) # v1.1.32
library(lmerTest) # v3.1.3
library(ggsci) # v3.0.0

# Set colorscales
npg_pal <- get_palette(palette = 'npg', 3)
npg_RL <- c(npg_pal[1], npg_pal[3])

# Read in data
data_AC <- read_csv('data_sup_fig01-2ce.csv')
data_BD <- read_csv('data_sup_fig01-2df.csv')
```

## Plot Panel A: Tau Burden vs. Overall Memory
```{r, echo = FALSE}
# PANEL A: Memory Overall vs. tau
figA <- data_AC %>% 
  ggplot(aes(x = ftp_braak34_suvr, y = dprime_total)) +
  geom_point() +
  geom_smooth(method = 'lm', limit = 0.9, linewidth = 2) +
  theme_classic() +
  xlab(bquote('[18F]FTP SUVR\n(Braak 3/4 Regions)')) +
  ylab('d\'') +
  theme(panel.background = element_blank(),
      axis.line = element_line(color = 'black'),
      axis.text.x = element_text(color = 'black', size = 8),
      axis.text.y = element_text(color = 'black', size = 8),
      axis.title.x = element_text(color = 'black', size = 10),
      axis.title.y = element_text(color = 'black', size = 10),
      aspect.ratio = 1)

figA
```

## Linear Mode: Tau Burden vs. Overall Memory
```{r, echo = FALSE}
print("+---------- Linear model ----------+")
dprime_mod <- lm(dprime_total ~ ftp_braak34_suvr + age + sex + yrs_edu,
                   data = data_AC)
summary(dprime_mod)
```

## Plot Panel B: Tau Burden vs. Valence-Specific Memory
```{r, echo = FALSE}
# PANEL B: Memory vs. tau by valence
figB <- data_BD %>% 
  ggplot(aes(x = ftp_braak34_suvr, y = dprime, color = valence)) +
  geom_point() +
  geom_smooth(method = 'lm', limit = 0.9, linewidth = 2) +
  theme_classic() +
  facet_wrap(.~valence) +
  scale_color_npg() +
  xlab(bquote('[18F]FTP SUVR\n(Braak 3/4 Regions)')) +
  ylab('d\'') +
  theme(panel.background = element_blank(),
      legend.position = 'none',
      axis.line = element_line(color = 'black'),
      axis.text.x = element_text(color = 'black', size = 8),
      axis.text.y = element_text(color = 'black', size = 8),
      axis.title.x = element_text(color = 'black', size = 10),
      axis.title.y = element_text(color = 'black', size = 10),
      aspect.ratio = 1)

figB
```

## Linear Mixed Effects Model: Tau Burden vs. Valence-Specific Memory
```{r, echo = FALSE}
# PANEL B: Stats
print("+---------- Linear Mixed Effects Model ----------+")
dprime_mod <- lmer(dprime ~ valence*ftp_braak34_suvr + age + sex + yrs_edu + (1 | subject), 
                   data_BD)
summary(dprime_mod)
```

## Plot Panel C: Tau Burden vs. Overall Learning Performance
```{r, echo = FALSE}
# PANEL C: Learning Overall vs. tau
figC <- data_AC %>% 
  ggplot(aes(x = ftp_braak34_suvr, y = learning_total)) +
  geom_point() +
  geom_smooth(method = 'lm', limit = 0.9, linewidth = 2) +
  theme_classic() +
  xlab(bquote('[18F]FTP SUVR\n(Braak 3/4 Regions)')) +
  ylab('Accuracy During Learning') +
  theme(panel.background = element_blank(),
      axis.line = element_line(color = 'black'),
      axis.text.x = element_text(color = 'black', size = 8),
      axis.text.y = element_text(color = 'black', size = 8),
      axis.title.x = element_text(color = 'black', size = 10),
      axis.title.y = element_text(color = 'black', size = 10),
      aspect.ratio = 1)

figC
```

## Linear Model: Tau Burden vs. Overall Learning Performance
```{r, echo = FALSE}
print("+---------- Linear model ----------+")
dprime_mod <- lm(learning_total ~ ftp_braak34_suvr + age + sex + yrs_edu,
                   data = data_AC)
summary(dprime_mod)
```

## Plot Panel D: Tau Burden vs. Valence-Specific Learning Performance
```{r, echo = FALSE}
# PANEL D: Learning vs. tau by valence
figD <- data_BD %>% 
  ggplot(aes(x = ftp_braak34_suvr, y = learning, color = valence)) +
  geom_point() +
  geom_smooth(method = 'lm', limit = 0.9, linewidth = 2) +
  theme_classic() +
  facet_wrap(.~valence) +
  scale_color_npg() +
  xlab(bquote('[18F]FTP SUVR\n(Braak 3/4 Regions)')) +
  ylab('Accuracy During Learning') +
  theme(panel.background = element_blank(),
      legend.position = 'none',
      axis.line = element_line(color = 'black'),
      axis.text.x = element_text(color = 'black', size = 8),
      axis.text.y = element_text(color = 'black', size = 8),
      axis.title.x = element_text(color = 'black', size = 10),
      axis.title.y = element_text(color = 'black', size = 10),
      aspect.ratio = 1)

figD
```

## Linear Mixed Effects Model: Tau Burden vs. Valence-Specific Learning Performance
```{r, echo = FALSE}
print("+---------- Linear Mixed Effects Model ----------+")
dprime_mod <- lmer(learning ~ valence*ftp_braak34_suvr + age + sex + yrs_edu + (1 | subject), 
                   data_BD)
summary(dprime_mod)
```

## Plot entire Figure s5
```{r, echo = FALSE}
behavioral_fig <- ggarrange(figA, figB, figC, figD,
          ncol = 2, nrow = 2,
          widths = c(0.33, 0.67),
          labels = c('C', 'D', 'E', 'F'))

ggsave(behavioral_fig, filename = 'sup_fig01-2cdef.pdf', width = 6.5, height = 6, units = 'in')

behavioral_fig
```