---
title: "Fig. S1 - Demographics & Participant Attrition"
author: "Tom Morin"
date: "2025-04-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse) # v2.0.0

df <- read_csv('data_sup_fig01-1a.csv')
```

## Count nubmer of Older Adults and Younger Adults
```{r, echo = FALSE}
# Count nubmer of Older Adults and Younger Adults
df %>% 
  group_by(age_group) %>%
  summarize(
    n = n()
)
```

## Demographics for older and younger adults with acceptable memory data
```{r, echo = FALSE}
# Demographics for older and younger adults with acceptable memory data
df %>%
  filter(exclude_memory == 0) %>%
  group_by(age_group) %>%
  summarize(
    n = n(),
    pct_female = sum(sex == 'F')/n,
    mean_age = mean(age, na.rm = T),
    sd_age = sd(age, na.rm = T),
    mean_yrs_edu = mean(yrs_edu, na.rm = T),
    sd_yrs_edu = sd(yrs_edu, na.rm = T)
  )
```

## Demographics for older and younger adults with acceptable fMRI data
```{r, echo = FALSE}
# Demographics for older and younger adults with acceptable fMRI data
df %>%
  filter(exclude_memory == 0, exclude_fmri == 0) %>%
  group_by(age_group) %>%
  summarize(
    n = n(),
    pct_female = sum(sex == 'F')/n,
    mean_age = mean(age, na.rm = T),
    sd_age = sd(age, na.rm = T),
    mean_yrs_edu = mean(yrs_edu, na.rm = T),
    sd_yrs_edu = sd(yrs_edu, na.rm = T)
  )
```

## Demographics for older and younger adults with acceptable FMT data
```{r, echo = FALSE}
# Demographics for older and younger adults with acceptable FMT data
df %>%
  filter(exclude_fmt == 0) %>%
  group_by(age_group) %>%
  summarize(
    n = n(),
    pct_female = sum(sex == 'F')/n,
    mean_age = mean(age, na.rm = T),
    sd_age = sd(age, na.rm = T),
    mean_yrs_edu = mean(yrs_edu, na.rm = T),
    sd_yrs_edu = sd(yrs_edu, na.rm = T)
  )
```

## Demographics for older adults with acceptable FTP data
```{r, echo = FALSE}
# Demographics for older adults with acceptable FTP data
df %>%
  filter(exclude_ftp == 0) %>%
  summarize(
    n = n(),
    pct_female = sum(sex == 'F')/n,
    mean_age = mean(age, na.rm = T),
    sd_age = sd(age, na.rm = T),
    mean_yrs_edu = mean(yrs_edu, na.rm = T),
    sd_yrs_edu = sd(yrs_edu, na.rm = T)
  )
```

## Demographics for older adults by PiB Status
```{r, echo = FALSE}
# Demographics for older adults by PiB Status
df <- df %>% mutate(pib_status = case_when(pib_dvr < 1.065 ~ 'PiB-',
                                           pib_dvr >= 1.065 ~ 'PiB+'))
  
df %>%  
  filter(exclude_memory == 0, exclude_pib == 0) %>%
  group_by(pib_status) %>%
  summarize(
    n = n(),
    pct_female = sum(sex == 'F')/n,
    mean_age = mean(age, na.rm = T),
    sd_age = sd(age, na.rm = T),
    mean_yrs_edu = mean(yrs_edu, na.rm = T),
    sd_yrs_edu = sd(yrs_edu, na.rm = T)
  )
```

## Demographics for older adults by Jack Status
```{r, echo = FALSE}
# Demographics for older adults by Jack Status
df <- df %>% mutate(tau_status = case_when(ftp_braak34_suvr < 1.26 ~ 'Tau-',
                                           ftp_braak34_suvr >= 1.26 ~ 'Tau+'),
                    jack_status = case_when(pib_status == 'PiB-' & tau_status == 'Tau-' ~ 'Healthy',
                                            pib_status == 'PiB-' & tau_status == 'Tau+' ~ 'PART',
                                            pib_status == 'PiB+' & tau_status == 'Tau-' ~ 'Jack A',
                                            pib_status == 'PiB+' & tau_status == 'Tau+' ~ 'Jack B',
                                            is.na(pib_status) & tau_status == 'Tau-' ~ 'Healthy',
                                            is.na(pib_status) & tau_status == 'Tau+' ~ 'PART',
                                            pib_status == 'PiB-' & is.na(tau_status) ~ 'Healthy',
                                            pib_status == 'PiB+' & is.na(tau_status) ~ 'Jack A'))
  
df %>%  
  filter(exclude_memory == 0, age_group == 'Older\nAdults') %>%
  group_by(jack_status) %>%
  summarize(
    n = n(),
    pct_female = sum(sex == 'F')/n,
    mean_age = mean(age, na.rm = T),
    sd_age = sd(age, na.rm = T),
    mean_yrs_edu = mean(yrs_edu, na.rm = T),
    sd_yrs_edu = sd(yrs_edu, na.rm = T)
  )

```

## Demographics for older adults by Biomarker Group
```{r, echo = FALSE}
# Demographics for older adults by Biomarker Group
df <- df %>% mutate(biomarker_group = case_when(jack_status == 'Healthy' ~ 'Normal Aging',
                                                jack_status == 'PART' ~ 'Normal Aging',
                                                jack_status == 'Jack A' ~ 'Preclinical AD',
                                                jack_status == 'Jack B' ~ 'Preclinical AD'))

df %>%
  filter(exclude_memory == 0, age_group == 'Older\nAdults') %>%
  group_by(biomarker_group) %>%
  summarize(
    n = n(),
    pct_female = sum(sex == 'F')/n,
    mean_age = mean(age, na.rm = T),
    sd_age = sd(age, na.rm = T),
    mean_yrs_edu = mean(yrs_edu, na.rm = T),
    sd_yrs_edu = sd(yrs_edu, na.rm = T)
  )

```