---
title: "Fig. 3: fMRI vs. Memory"
author: "Tom Morin"
date: "2025-12-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = FALSE}
# Load Libraries
library(tidyverse) # v2.0.0
library(ggpubr) # v0.6.0
library(lme4) # v1.1.32
library(lmerTest) # v3.1.3
library(rstatix) # v0.7.2
library(lsmeans) #2.30.2
library(sjPlot) #v2.9.0

# Set colorscales
biomarker_group_cscale <- scale_color_manual(values = c("#549e3e", "#323690", "#B6357B"))
biomarker_group_fscale <- scale_fill_manual(values = c("#549e3e", "#323690", "#B6357B"))

# Read data
data_BC <- read_csv('data_fig3bc.csv')

data_BC$biomarker_group <- factor(data_BC$biomarker_group,
                                  levels = c('Younger\nAdults', 'Normal\nAging',
                                             'Preclinical\nAD'))

data_D <- read_csv('data_fig3d.csv')

data_D$biomarker_group <- factor(data_D$biomarker_group,
                                  levels = c('Younger\nAdults', 'Normal\nAging',
                                             'Preclinical\nAD'))

data_D <- data_D %>% mutate(biomarker_group_int = case_when(biomarker_group == 'Younger\nAdults' ~ 0,
                                                            biomarker_group == 'Normal\nAgers' ~ 1,
                                                            biomarker_group == 'Preclinical\nAD' ~ 2))
```

## Plot Panel B: Remembered vs. Forgotten BOLD signal contrast, across biomarker groups
```{r, echo = FALSE}
# Panel B
within_ttests <- data_BC %>%
  group_by(biomarker_group) %>%
  summarise(P = t.test(bold, mu = 0)$p.value,
            Sig = ifelse(P < 0.05, "*", ""),
            MaxWidth = max(bold))

figB <- ggbarplot(data_BC, 
                   x = "biomarker_group", 
                   y = "bold", 
                   fill = "biomarker_group", 
                   alpha = 0.5,
                   palette = c("#549e3e", "#323690", "#B6357B"), 
                   add = "mean_se", 
                   position = position_dodge(0.8),
                   xlab = '', 
                   ylab = 'fMRI Beta Coefficient\n(Remembered vs. Forgotten)', 
                   legend = 'none',
                   order = c('Younger\nAdults', 'Normal\nAging', 
                             'Preclinical\nAD')) +
  geom_hline(yintercept = 0) +
  geom_jitter(aes(color = biomarker_group, fill = biomarker_group), alpha = 0.5,
              position = position_jitterdodge(0.1)) +
  theme(panel.background = element_blank(),
      axis.line = element_line(color = 'black'),
      legend.position = 'none', 
      axis.text.x = element_text(color = 'black', size = 8),
      axis.text.y = element_text(color = 'black', size = 8),
      axis.title.x = element_text(color = 'black', size = 10),
      axis.title.y = element_text(color = 'black', size = 10),
      legend.title = element_blank(),
      aspect.ratio = 1)

group.stat.test <- data_BC %>% 
  t_test(bold ~ biomarker_group, detailed = TRUE, paired = FALSE)
p.group.stat.test <- group.stat.test %>%
  filter(p < 0.05) %>%
  add_xy_position(x = "biomarker_group", fun = "mean_se", dodge = 0.8)
p.group.stat.test <- p.group.stat.test %>% mutate(p.signif = ifelse(p < 0.05, "*", ""))

figB <- figB + stat_pvalue_manual(p.group.stat.test, 
                                  label = 'p.signif', 
                                  tip.length = 0.01,
                                  step.increase = 0.02,
                                  y.position = 0.4,
                                  size = 6)
figB
```

## Are there group differences in parahippocampal activation?
```{r, echo = FALSE}
# Stats for Panel A

# All subjects, simple model
lm1 <- lm(bold ~ biomarker_group, data = data_BC)
print('==========================================================================')
summary(lm1)
print('==========================================================================')
```

## Post-hoc T-tests
```{r, echo = FALSE}
group.stat.test <- data_BC %>% 
  t_test(bold ~ biomarker_group, detailed = TRUE, paired = FALSE,
         p.adjust.method = 'none')

group.stat.test
```

## Repeat analysis, controlling for learning performance
```{r, echo = FALSE}
# All subjects: memory, learning regressed
lm2 <- lm(bold ~ biomarker_group + total_learning, data = data_BC)
print('==========================================================================')
summary(lm2)
print('==========================================================================')

data_BC <- data_BC %>% mutate(biomarker_group_int = case_when(biomarker_group == 'Younger\nAdults' ~ -1,
                                                              biomarker_group == 'Normal\nAging' ~ 0,
                                                              biomarker_group == 'Preclinical\nAD' ~ 1))


lm1 <- lm(bold ~ biomarker_group_int, data = data_BC)
lm2 <- lm(bold ~ biomarker_group_int*dprime_total, data = data_BC)
lm3 <- lm(bold ~ biomarker_group_int*total_learning, data = data_BC)
lm4 <- lm(bold ~ biomarker_group_int*dprime_total + biomarker_group_int*total_learning, data = data_BC)

tab_model(lm1, lm2, lm4)
```

## Is RvF BOLD signal contrast related to memory performance?
```{r, echo = FALSE}
modA <- lm(dprime_total ~ bold*biomarker_group_int + sex, data = data_BC)
summary(modA)

slopes <- emtrends(modA, ~ biomarker_group_int, var = 'bold')
summary(slopes, infer=TRUE)

modB <- lm(dprime_total ~ bold*biomarker_group_int + sex + total_learning, data = data_BC)
summary(modB)

slopes <- lstrends(modB, ~ biomarker_group_int, var = 'bold')
summary(slopes, infer=TRUE)

figD <- data_BC %>%
  ggplot(aes(x = bold, y = dprime_total, group = biomarker_group, 
             color = biomarker_group, fill=biomarker_group)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = 'lm', level = 0.9, linewidth = 2) +
  biomarker_group_cscale +
  biomarker_group_fscale +
  theme_classic() +
  xlab(bquote('fMRI Beta Coefficient\n(Remembered vs. Forgotten)')) +
  ylab('Memory (d\')') +
  theme(panel.background = element_blank(),
      axis.line = element_line(color = 'black'),
      legend.position = 'none', 
      axis.text.x = element_text(color = 'black', size = 8),
      axis.text.y = element_text(color = 'black', size = 8),
      axis.title.x = element_text(color = 'black', size = 10),
      axis.title.y = element_text(color = 'black', size = 10),
      legend.title = element_blank(),
      aspect.ratio = 1)
figD

tab_model(modA, modB)
```

## Plot Panel C: FMT vs BOLD contrast, by biomarker group
```{r, echo = FALSE}
# Panel C
figC <- data_BC %>%
  ggplot(aes(x = fmt_ki_vst, y = bold, group = biomarker_group, 
             color = biomarker_group, fill=biomarker_group)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = 'lm', level = 0.9, linewidth = 2) +
  biomarker_group_cscale +
  biomarker_group_fscale +
  theme_classic() +
  xlab(bquote('[18F]FMT Ki\n(Ventral Striatum)')) +
  ylab('fMRI Beta Coefficient\n(Remembered vs. Forgotten)') +
  theme(panel.background = element_blank(),
      axis.line = element_line(color = 'black'),
      legend.position = 'none', 
      axis.text.x = element_text(color = 'black', size = 8),
      axis.text.y = element_text(color = 'black', size = 8),
      axis.title.x = element_text(color = 'black', size = 10),
      axis.title.y = element_text(color = 'black', size = 10),
      legend.title = element_blank(),
      aspect.ratio = 1)

figC
```

## Linear Model: Does Biomarker Group mediate the relationship between FMT and BOLD signal?
```{r, echo = FALSE}
# Stats for Panel B
# All subjects, simple model
lm1 <- lm(bold ~ biomarker_group*fmt_ki_vst, data = data_BC)
print('==========================================================================')
summary(lm1)
print('==========================================================================')
```

## Repeat analysis, controling for learning performance
```{r, echo = FALSE}
# All subjects: memory, learning regressed
lm2 <- lm(bold ~ biomarker_group*fmt_ki_vst + total_learning, data = data_BC)
print('==========================================================================')
summary(lm2)
print('==========================================================================')
```

## Plot entire Figure 3bcd
```{r, echo = FALSE}
fmri_mem <- ggarrange(figB, figD, figC, 
          ncol = 3, 
          labels = c('B', 'C', 'D'),
          align = 'v')

ggsave(fmri_mem, filename = 'fig3.pdf', width = 6, height = 3, units = 'in')
fmri_mem
```
