---
title: "Fig S7 - Positivity Effect and Tau Burden"
author: "Tom Morin"
date: "2025-04-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load Libraries
library(tidyverse) # v2.0.0
library(ggpubr) # v0.6.0
library(rstatix) # v0.7.2
library(lme4) # v1.1.32
library(lmerTest) # v3.1.3

# Set colorscales
jack_status_cscale <- scale_color_manual(values = c("#549e3e", "#323690", "#B6357B"))
jack_status_fscale <- scale_fill_manual(values = c("#549e3e", "#323690", "#B6357B"))

df <- read_csv('data_sup_fig07.csv')

df$tau_status <- factor(df$tau_status, levels = c('Younger\nAdults', 
                                                  'Tau-', 'Tau+'))
```

## Plot Panel A: Positivity Index vs. Tau Burden
```{r, echo = FALSE}
figA <- df %>% 
  ggplot(aes(x = ftp_braak34_suvr, y = pos_index)) +
  geom_point() +
  geom_smooth(method = 'lm', limit = 0.9, linewidth = 2) +
  theme_classic() +
  xlab(bquote('[18F]FTP SUVR\n(Braak III/IV Regions)')) +
  ylab(bquote(atop('Positivity Index', (H[Rew] - H[Loss])/(H[Rew] + H[Loss])))) +
  theme(panel.background = element_blank(),
      axis.line = element_line(color = 'black'),
      axis.text.x = element_text(color = 'black', size = 8),
      axis.text.y = element_text(color = 'black', size = 8),
      axis.title.x = element_text(color = 'black', size = 10),
      axis.title.y = element_text(color = 'black', size = 10),
      aspect.ratio = 1)

figA
```

## Linear Model to test for relationship between tau burden and positivity index
```{r, echo = FALSE}
# Stats for Panel A

# Simplest model
lm1 <- lm(pos_index ~ ftp_braak34_suvr + age + sex + yrs_edu, data = df)
summary(lm1)
```

## Repeat analysis while accoutning for learning performance
```{r, echo = FALSE}
# Model accounting for learning accuracy
lm2 <- lm(pos_index ~ ftp_braak34_suvr + learning_RewVsLoss, data = df)
summary(lm2)

```

## Plot Panel B: FMT vs. Positivity Index, by Tau Burden
```{r, echo = FALSE}
figB <- df %>% filter(!is.na(tau_status)) %>%
  ggplot(aes(x = fmt_ki_vst, y = pos_index, group = tau_status,
             color = tau_status, fill = tau_status)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = 'lm', limit = 0.9, linewidth = 2) +
  scale_fill_manual('[18F]FTP Status', values = c("#549e3e", "#323690", "#B6357B")) +
  scale_color_manual('[18F]FTP Status', values = c("#549e3e", "#323690", "#B6357B")) +
  theme_classic() +
  xlab(bquote('[18F]FMT Ki\n(Ventral Striatum)')) +
  ylab('\u0394 d\' (Reward vs. Loss)') +
  theme(panel.background = element_blank(),
      axis.line = element_line(color = 'black'),
      axis.text.x = element_text(color = 'black', size = 8),
      axis.text.y = element_text(color = 'black', size = 8),
      axis.title.x = element_text(color = 'black', size = 10),
      axis.title.y = element_text(color = 'black', size = 10),
      aspect.ratio = 1)

figB
```

## Linear Model: Interaction between FMT and Tau Burden on Positivity index?
```{r, echo = FALSE}
# Stats for Panel B
lm1 <- lm(pos_index ~ fmt_ki_vst*ftp_braak34_suvr + age + sex + yrs_edu, 
          data = df %>% filter(!is.na(ftp_braak34_suvr)))
summary(lm1)
```

## Repeat analysis, accounting for learning performance
```{r, echo = FALSE}
# Model accounting for learning accuracy
lm2 <- lm(pos_index ~ fmt_ki_vst*ftp_braak34_suvr + age + sex + yrs_edu + learning_RewVsLoss, 
          data = df %>% filter(!is.na(ftp_braak34_suvr)))
summary(lm2)

```


```{r, echo = FALSE}
pos_fig <- ggarrange(figA, figB, 
          ncol = 2, 
          labels = c('A', 'B'),
          widths = c(0.6, 1))

ggsave(pos_fig, filename = 'sup_fig07.pdf', width = 6, height = 3, units = 'in')
pos_fig
```